NMRlipids Databank: Making data-driven analyses of membrane properties accessible for all

### 2022-11-22 main.tex
Updates from Overleaf

\title{NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}

@@

While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, researches are now searching new avenues by building models not only for individual molecules but whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.

The importance of sharing MD simulation data following the FAIR principles~\cite{wilkinson16} has been widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20}, and databanks are emerging~\cite{meyer10,kamp10,hospital16,mixcoha16,newport19,bekker20,espigares20,leston22}.
The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such quality evaluation has in some cases also been implemented~\cite{meyer10,hospital16}. However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging. 
While importance of such databanks for MD simulations is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and different kinds of approaches are emerging~\cite{meyer10,kamp10,hospital16,mixcoha16,newport19,bekker20,espigares20,leston22}, generally accepted protocols and best practices are still under active development.

Here we present a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.  The concept of overlay databank is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it can be  potentially used for wide range of situations. Particularly when storage of raw data requires significant resources and final outcomes or best practices are not yet clear, overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.

The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data, data-driven analysis detecting correlations between properties of model cell membranes and analyses of rare phenomena that are beyond the scope of standard MD simulation studies. The NMRlipids databank provides new tools for researchers in wide range of fields in academia and industry from cell membrane biology to lipid nanoparticle formulations and data-driven computational chemistry and machine learning.

@@

%Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
\subsection{Programmatic access to MD simulation data of membranes composed of most common biologically abundant lipids}
NMRlipids databank is a community driven databank containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}. Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using Python API or graphical user interface (\url{http://www.databank.nmrlipids.fi/}). NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content. Following the FAIR principles and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}. The NMRlipids databank will be always open for data and other contributions.

The distribution of different lipid types in simulations in the NMRlipids databank (Fig.~\ref{fig:overlay}C) resembles the biological abundancies, the PC being the most common lipid type followed by PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}. 
Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}. The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.

The three layer structure of the NMRlipids databank, illustrated in Fig.~\ref{fig:overlay} a, enables an efficient way to share and upcycle large MD simulation trajectories.  Layer 1 contains raw simulation data that can locate in any publicly available repository or server offering long term stability and permanent links to the data, such as digital object identifiers. Layer 2 contains all the relevant information on the simulations, including links to the raw data in layer 1 and relevant metadata describing the systems. Layer 3 is the application layer containing outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses desribed in sections below.

Layer 2 is the core of the databank. In addition to the metadata and links to raw simulation data, it contains universal naming conventions for the molecules and atoms in the databank, computer programs to generate and analyse the databank, basic properties calculated from all simulations (area per lipid, C-H bond order parameters, x-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data. In practise, this information is stored is the git repository which is currently available at~\url{https://github.com/NMRLipids/Databank}.
Detailed description on its content is in the supplementary information. Applications in layer 3 uses the information in layer 2 to access raw data in layer 1 to perform automatic analyses over large sets of simulations in the databank.

@@

\begin{figure}
    \centering
    \includegraphics[width = 180mm]{Figures/overlay2.pdf}
    \caption{a) Structure of an overlay databank. 
    %Raw data is available from a publicly accessible repository (layer 1).
    %The core of the databank contains information on the location of the raw data and indexing of molecular %compositions and simulation details described with universal naming conventions (layer 2).
    %The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3). Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
    d) Currently available binary mixtures in the NMRlipids databank. 
    e) Distribution of force fields in the simulations in the NMRlipids databank.
    The figures and numbers are created on 9th of May 2022.}
    \label{fig:overlay}
\end{figure}

\subsection{Quality evaluation of force fields}
Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19}, but universal quality measure for membrane simulations is not defined and controversial results are often reported from simulations~\cite{antila22b}. The lack of universal quality measure for membrane simulations complicates the selection of proper simulation parameters and the estimation of reliability in simulations, thereby being a major obstacle in many applications of membrane MD simulations. To enable the rapid quality evaluation of membrane MD simulations, a quantitative quality measure based on C-H bond order parameters from NMR experiments is defined in the NMRlipids databank. %Conformational ensembles of individual lipid molecules are evaluated against C-H bond order parameters from NMR experiments. 
This is complemented by comparing the locations of x-ray scattering form factor minima against experiments. 
%that are related to the overall membrane dimensions via electron density profile. 
Both of these are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}. The used quality measures are defined in detail in the supplementary information.

%NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications. 
Simulations with the highest scores based on order parameter evaluation for all simulations are shown in Fig.~\ref{fig:quality} A and for POPE lipids in Fig.~\ref{fig:quality} B. The direct comparison with experiments for the best simulation, POPC with OPLS3e, is shown in Fig.~\ref{fig:quality} C, where discrepancies with experiments are observed only for the third carbon in the $\it{sn}$-1 chain, and carbons 8 and 9 close to the double bond in $\it{sn}$-2 chain. The best overall quality for POPE is found in GROMOS-CKP simulation, for which the direct comparison in Fig.~\ref{fig:quality} D show minor differences in acyl chains, but major discrepancies in glycerol backbone and in the first carbon of $\it{sn}$-2 chain. These parts are better described for POPE by CHARMM36 parameters (direct comparison shown in Fig.~\ref{fig:quality} E), but its overall quality suffers from the overestimated order in acyl chains.

\begin{figure}[tbp]
    \centering
    \includegraphics[width=180mm]{Figures/quality3.pdf}
    \caption{ A) Best simulations ranked based on overall order parameter quality.
    B) Best simulations ranked based on the overall order parameter quality of POPE lipid. 
    C) Scatter plots and Pearson correlation coefficients for the area per lipid, thickness, and experimental observables. 
    %second minima of form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank. 
    All correlation coefficients have p-value below 0.001. For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
    D)-F) Evaluation against experimental data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
    %c) Change in area per lipid upon addition of POPE to POPC in different simulation extracted from the NMRlipids databank and quality evaluation table of POPE simulations.
    }
    \label{fig:quality}
\end{figure}

\subsection{Finding the best models for PC and PE lipid mixtures}
The power of the NMRlipids databank in finding the best models for a certain lipid bilayer system is exemplified here for mixtures of PE and PC lipids. Area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K are shown in Fig.~\ref{fig:POPCPOPEapls}. Because the area per lipid strongly correlates with the average order of $\it{sn}$-1 chain (Fig.~\ref{fig:quality} F), we use order  parameter qualities to evaluate which simulations predict the best area per lipids for POPC and POPE. Among the force fields with area per lipids shown in Fig.~\ref{fig:POPCPOPEapls}, Slipids ranks 1st for POPC with a clear difference to others, and 2nd for POPE with only marginally lower quality than GROMOS-CKP (Figs.~\ref{fig:quality} A and B). The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that GROMOS-CKP and CHARMM36 predicts too ordered (too low area per lipid) bilayer for POPC, while the order parameters in Lipid17 are lower than in experiments indicating overestimated area per lipid. For POPE, order parameters in  Slipids and GROMOS-CKP are close to experiments suggesting that their are per lipids are most reasonable, while CHARMM36 and Lipid17 are too packed thereby overestimating the order parameters. 
\begin{figure}[tbp]
    \centering
    \includegraphics[width=88mm]{Figures/APL.pdf}
    \caption{Differences in area per lipids of POPC and POPE bilayers between different force fields.  
    }
    \label{fig:POPCPOPEapls}
\end{figure}

In conclusion, the quality evaluation based on the NMRlipids databank suggests that the Slipids parameters would give the most reliable description for membrane packing of POPC:POPE mixtures. Slipids was a relatively good model also for POPC:POPS mixtures in similar comparisons utilizing the preliminary version of the NMRlipids databank~\cite{antila22b}, although charged membranes are more complicated as the counterion binding affinity plays a significant role in membrane packing~\cite{antila19}. On the other hand, CHARMM36 overcomes Slipids when the focus is on conformational ensembles of headgroups  in different types of lipids because the accuracy of glycerol backbone is low in Slipids~\cite{bacle21}. Altogether, these examples demonstrate how the NMRlipids databank can be used to navigate in the complex landscape of lipid force field quality.

@@

\subsection{Water diffusion anisotropy in membrane systems}
The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}. Several models on how water dynamics depends on membrane properties have been proposed based on different experimental and theoretical methods, yet consensus 
%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
has not been reached~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}. MD simulations have been also used in these endeavours, but systematic studies are limited by the accessible amount of data as only few water permeation events are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}. Therefore, this serves as an excellent example where the large amount of available MD simulations in the NMRlipids databank can be utilized to answer important biological questions.

To this end, we first calculated the water permeability through membranes from all the simulations in the NMRlipids databank. Resulting permeabilities are shown as a function of temperature, membrane thickness, area per lipid, and acyl chain order in Figs.~\ref{fig:permeability} B-E. The values for water permeabilities from simulations vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively. These values have the same order of magnitude as the experimentally determined diffusive permeability coefficients, but are on average below the values reported for PC lipids in liquid crystalline phase,~190-330\,$\mu$m/s~\cite{jansen95}. To analyze how water permeability through membranes depend on bilayer properties on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range. These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots. As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B. Water permeability decreases with increasing membrane packing, i.e., with decreasing area per lipid and increasing thickness and acyl chain order. Linear dependence of permeability on area per lipid and thickness is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. Our results suggest on average linear dependence for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D), yet stronger dependence is observed for more loosely packed membranes. 
%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
%, although weak decrease for the latter two may be visible.

@@

Applications of MD simulations in understanding biomolecular behaviour are steadily increasing, but protocols for quality evaluation and sharing simulation data fall behind the ones in more elaborated fields, such as structural biology. Development of such protocols has been limited by practical challenges, such as lack of universally defined quality measures for MD simulations, incompatible naming conventions in different models and simulation softwares, and high technical requirements for hardware to share large trajectories from MD simulations, as well as incentives for scientist to evaluate and share their MD simulation data.

>>>>

\title{NMRlipids Databank: Making data-driven analyses of membrane properties accessible for all}
%Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}

@@

Cellular membranes are composed of hundreds of different types of lipid molecules which regulate membrane properties, morphology and biological functions~\cite{vanmeer08}. Membrane lipid composition can be also related to diseases, such as cancer and neurodegenerative disorders, and potential therapeutics that affect membrane compositions have been proposed~\cite{torres21}. However, membrane containing systems are often difficult to study experimentally because membranes are in disordered fluid state in biological conditions and sample preparation can be highly complicated due to rich phase behaviour and complex interactions between lipids and other biomolecules. Therefore, connections between complex lipid interactions and biological functions are still poorly understood.
%why such variety of lipids is needed and how their. 
Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, yet their accuracy is often compromised by the quality of model parameters and other artefacts~\cite{antila22b,gupta22}. 
%While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, 
On the  other hand, the accuracy of models is becoming increasingly important as researches are progressing from the simulations of 
%now searching new avenues by building models not only 
individual molecules to the models of whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}. In such systems exhibiting very complex behaviour, inaccuracies are more difficult to detect and multiplication of small errors may have major effects the conclusions.

In contrast to experimental structural biology where standard protocols to share and quality evaluate the resolved structures are established in the PDB databank~\cite{montelione13}, such best practices are yet to be defined for MD simulations. Importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and solutions for sharing data are emerging for proteins in solution~\cite{meyer10,kamp10} or membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, and arbitrary biomolecules~\cite{bekker20}. However,
%The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such 
automatic quality evaluation~\cite{meyer10,hospital16} and programmatic access are still rare. 
%However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging. 
Therefore, tools for automatic quality evaluation of membrane simulations or training sets for machine learning models of membrane containing systems are not yet available.

%Applications of MD simulations in understanding biomolecular behaviour are steadily increasing, but protocols for quality evaluation and sharing simulation data fall behind the ones in more elaborated fields, such as structural biology. Development of such protocols has been limited by practical challenges, such as lack of universally defined quality measures for MD simulations, incompatible naming conventions in different models and simulation softwares, and high technical requirements for hardware to share large trajectories from MD simulations, as well as incentives for scientist to evaluate and share their MD simulation data.

Here we present the NMRlipids databank, a community driven open access databank with programmatic access containing atomistic resolution MD simulations of lipid bilayers. The programmatic access enables various data-driven approaches, providing new tools for researchers in wide range of fields in academia and industry from cell membrane biology to lipid nanoparticle formulations and data-driven computational chemistry and machine learning. Here we demonstrate how data-driven analysis of water anisotropic diffusion in membrane systems from the NMRlipids databank can extend the scope of MD simulations to MRI imaging and pharmacokinetics, and how the NMRlipids databank can be utilized to analyse rare phenomena that are beyond the scope of standard MD simulation studies. Furthermore, we perform automatic quality evaluation of membrane simulations which guides the selection of best models for specific applications and development of simulation parameters and methodology.
%The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data,

While the NMRlipids databank currently contains only bilayer systems, it can be readily expanded to contain other molecules than lipids, such as disordered proteins or sugars. 
%a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.  
The combination of overlay databank structure and open collaboration introduced here can be applied also when building databanks in other fields,  
%is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it can be  potentially used for wide range of situations. 
particularly when storage of raw data requires significant resources, best practices in the field are not defined, and incentives to share data do not exist.
%overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
Therefore, the approach opens up possibilities to create databanks enabling data-driven and machine learning applications in new fields.

@@

%Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
\subsection{Programmatic access to MD simulation data of membranes composed of most common biologically abundant lipids}
NMRlipids databank is a community driven databank containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}. Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using Python API or graphical user interface (\url{http://www.databank.nmrlipids.fi/}). 
%NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content. 
Following the FAIR principles~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}. 
%The NMRlipids databank will be always open for data and other contributions.

Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D. While these compose the majority of available trajectories, also mixtures up to five lipids are available. The distribution of available simulations containing a specific lipid is shown in Fig.~\ref{fig:overlay}C. This roughly resembles the biological abundancies of different lipid types, PC being the most common  followed by cholesterol, PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}. The force fields used in the simulations cover all the essential parameters used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields. Therefore, the averages calculated over the databank can be considered as mean predictions from available lipid models for an average cell membrane.
%Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}. The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.

The efficient upcycling of large MD simulation trajectories in the NMRlipids databank is enabled by the overlay structure illustrated in Fig.~\ref{fig:overlay}A. Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability and permanent links to the data, such as digital object identifiers. The {\it Databank layer} is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, computer programs to create and analyse the entries, basic properties calculated from all simulations (area per lipid, C-H bond order parameters, x-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data. In practise, this information is stored is the git repository which is currently available at~\url{https://github.com/NMRLipids/Databank}. The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses described in sections below. More detailed description on the NMRlipids databank structure is available the supplementary information.

@@

\begin{figure}
    \centering
    \includegraphics[width = 180mm]{Figures/overlay2.pdf}
    \caption{a) Structure of an overlay databank. 
    %Raw data is available from a publicly accessible repository (layer 1).
    %The core of the databank contains information on the location of the raw data and indexing of molecular %compositions and simulation details described with universal naming conventions (layer 2).
    %The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3). Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
    d) Currently available binary mixtures in the NMRlipids databank. 
    e) Distribution of force fields in the simulations in the NMRlipids databank.
    }
    \label{fig:overlay}
\end{figure}

\subsection{Selection of best simulation parameters for applications using the NMRlipids databank}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
Understanding the accuracy of used force field parameters is crucial when estimating the reliability and significance MD simulation results. However, the lack of universal quality measures for MD simulations and complicated landscape of force field quality for membranes hamper the quality of simulations which has resulted to many controversial results~\cite{antila22b}. For example, previous quality evaluations of membrane simulations have concluded that CHARMM36 parameters give the best description for lipid headgroup conformational ensembles~\cite{bacle21}, GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}, and that OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. 
%, thereby being a major obstacle in many applications of membrane MD simulations. 
To enable the navigation in this complex force field quality landscape, we define here quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide force field development. For detailed definitions of quality measures, see the supplementary information.

\begin{figure}[!t]
    \centering
    \includegraphics[width=180mm]{Figures/quality3.pdf}
    \caption{ A) Best simulations ranked based on overall order parameter quality.
    B) Best simulations ranked based on the overall order parameter quality of POPE lipid. 
    C)-E) Evaluation against experimental data exemplified for a simulation with the best overall order parameter quality (C), the best quality for POPE lipid (D), and the headgroup quality for POPE (E).
    F) Scatter plots and Pearson correlation coefficients for the membrane area per lipid, thickness, first minima of x-ray scattering form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank. 
    All correlation coefficients have p-value below 0.001. For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
    }
    \label{fig:quality}
\end{figure}

Ranking of all simulations with the highest scores based on evaluation against C-H bond order parameters from NMR experiments is shown in Fig.~\ref{fig:quality} A and ranking of best POPE lipid bilayer simulations in Fig.~\ref{fig:quality} B. Direct comparisons with experiments for selected simulations are shown in Figs.~\ref{fig:quality} C-D. C-H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, but they correlate well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality} F. This is complemented by comparing the absolute values of x-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions via electron density profile. Particularly, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality} F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.

%best simulation, POPC with OPLS3e, is shown in
%based on C-H bond order parameters from NMR experiments and x-ray scattering form factors that enable the rapid quality evaluation of membrane MD simulations in the NMRlipids databank.

%Order parameters and scattering form factors are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.

%NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.

%Fig.~\ref{fig:quality} C, where discrepancies with experiments are observed only for the third carbon in the $\it{sn}$-1 chain, and carbons 8 and 9 close to the double bond in $\it{sn}$-2 chain. The best overall quality for POPE is found in GROMOS-CKP simulation, for which the direct comparison in Fig.~\ref{fig:quality} D show minor differences in acyl chains, but major discrepancies in glycerol backbone and in the first carbon of $\it{sn}$-2 chain. These parts are better described for POPE by CHARMM36 parameters (direct comparison shown in Fig.~\ref{fig:quality} E), but its overall quality suffers from the overestimated order in acyl chains.

%\subsection{Finding the best models for PC and PE lipid mixtures}
The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K. Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE. Because the area per lipid correlates with the average order parameter of $\it{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation. Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1st for POPC with a clear difference to others, and 2nd for POPE with only marginally lower quality than GROMOS-CKP, therefore being the best selection for studies where packing effects of PE lipids are relevant. 
%The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that 
Simulations with CHARMM36 and GROMOS-CKP predict too packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while area per lipid for POPC is overstimated in Lipid17. For POPE, 
%order parameters in  Slipids and GROMOS-CKP are close to experiments suggesting that their are per lipids are most reasonable, while
CHARMM36 and Lipid17 predict too packed membranes.

In conclusion, 
%the quality evaluation based on the NMRlipids databank suggests that the Slipids parameters would give the most reliable description for membrane packing of POPC:POPE mixtures. Slipids was a relatively good model also for POPC:POPS mixtures in similar comparisons utilizing the preliminary version of the NMRlipids databank~\cite{antila22b}, although charged membranes are more complicated as the counterion binding affinity plays a significant role in membrane packing~\cite{antila19}. On the other hand, CHARMM36 overcomes Slipids when the focus is on conformational ensembles of headgroups  in different types of lipids because the accuracy of glycerol backbone is low in Slipids~\cite{bacle21}. Altogether, these examples demonstrate how 
the automatic quality evaluation in the NMRlipids databank enables rapid selection of the best simulation parameters to specific applications without extensive and tedious manual force field evaluation. This possibility will promote more reliable simulation results for wide range of applications, support force field development and parametrization of coarse grained force fields against the best atomistic MD simulations. The automatically extracted quantitative quality measures will be particularly useful to guide automatic parametrization procedures.

\begin{figure}[tbp]
    \centering
    \includegraphics[width=88mm]{Figures/APL.pdf}
    \caption{Differences in area per lipids of POPC and POPE bilayers between different force fields.  
    }
    \label{fig:POPCPOPEapls}
\end{figure}

@@

\subsection{Detection rare phenomena using NMRlipids databank: Cholesterol flip-flops}
Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}. Phospholipid flip-flops are slow, occuring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range between minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}. These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip-flop events have been published recently~\cite{gu19,javanainen19,baral20}. 
%However, systematic studies on how cholesterol flip-flop rate depends on membrane properties are still beyond the current computational resources available for a single research group. 
These studies report increase in cholesterol flip-flop rates with increasing unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip-flop rates and membrane properties have not been systematically studied. Here we demonstrate that the large amount of simulation data available in the NMRlipids databank with programmatic access makes the analysis of these correlations accessible for all.

Flip-flops were observed for cholesterol, % (634 events in 77 different simulations), 
DCHOL (18,19-di-nor-cholesterol), %, 16 flip-flops in 1 simulation) 
DOG (1,2-dioleoyl-sn-glycerol), %, 44 flip-flops in 4 simulations) 
and SDG (1-stearoyl-2-docosahexaenoyl-sn-glycerol) in the NMRlipids databank. Cholesterol flip-flop rates range between 0.001-1.6\,$\mu$s$^{-1}$ with the mean of 0.16\,$\mu$s$^{-1}$ and median of 0.07\,$\mu$s$^{-1}$. These values are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}. Flip-flop rates of DCHOL, 0.2\,$\mu$s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,$\mu$s$^{-1}$ and 0.5\,$\mu$s$^{-1}$, respectively. Flip-flops were not observed for other lipids, giving the upper limits for PC lipid flip-flop rate of 9\,s$^{-1}$ and for ceramide (N-palmitoyl-D-erythro-sphingosine) of 0.002\,$\mu$s$^{-1}$. Therefore, data in the NMRlipids databank suggests that the lipid flip-flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides. However, amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.

Nevertheless, we use the wide range of available simulations to analyse how cholesterol flip-flop rate depends on membrane properties. Cholesterol flip-flop rates and their averages over fixed ranges of x-axis values are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B-D. The results reveal a non-linear correlation between decreasing cholesterol flip-flop rate and membrane packing. Flip-flop rates increase with an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing. The order of magnitude changes in cholesterol flip-flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}. Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip-flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.

\begin{figure}[htb]
    \centering
    \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
    \caption{{\bf A} Illustration of cholesterol flip-flop.  
      {\bf B-D} Cholesterol flip-flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order. Values from simulations with non-zero flip-flop rates are shown with blue dots. Averages over fixed range of x-axis values are shown with black dots.
    }
    \label{fig:flip-flops}
\end{figure}

Cholesterol flip-flops are phenomena whose analyses from MD simulations has only recently became possible for groups with substantial resources and special expertise available~\cite{gu19,javanainen19,baral20}. Our analysis demonstrates how the NMRlipids databank makes analyses of such rare phenomena accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data.

\subsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}. MD simulations are rarely used to analyze anistropic diffusion of water as only few permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount data challenging. Here we show that the NMRlipids databank can be used to analyze how anisotropic water diffusion in a multilamellar membrane systems depends on membrane properties.

@@

The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}. On the other hand, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}. The NMRlipids databank will support the development in all these directions. The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the multiplication of small errors in large simulations. On the other hand, the NMRlipids databank serves as a training set for diverse machine learning applications. 
%Most straightforward applications include models that would predict connections between membrane properties that are already stored in the databank. 
For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments. On the other hand, more elaborated models can be trained to predict arbitrary membrane properties using the data from the NMRlipids databank.
